---
title: "Boston Nonlinear Dataset"
output: html_notebook
---

This is Problem 8 of ISLR's Chapter 7 Applied Section

This question uses the variables dis (the weighted mean of distances
to five Boston employment centers) and nox (nitrogen oxides concentration in parts per 10 million) from the Boston data. We will treat
dis as the predictor and nox as the response.

Very low p-values for all the degrees of the polynomial that our cubic fit is a good fit, but we might need to increase the flexibility of our model

By evaluating the polynomial RSS, we see that things start to level off after third degree.
```{r}
library(tidyverse)
library(ISLR)
library(MASS)

#Part A: Fit a cubic polynomial to predict nox from dis
model = glm(nox~poly(dis,3), data = Boston)
summary(model)

dis.grid = seq(min(Boston$dis),max(Boston$dis), by = .1)
preds = predict(model,list(dis = dis.grid))

plot(nox~dis, data = Boston, col = "darkgrey")
lines(dis.grid,preds,lwd = 2, col = "red")

#Part B: Determining the best degree polynomial based off of the residuals
RSS = rep(NA,10)
for (i in 1:10) {
  model = glm(nox~poly(dis,i), data = Boston)
  RSS[i] = mean((model$residuals)^2)
}
ggplot() + 
  geom_smooth(mapping = aes(x = 1:10,y = RSS))

```

Now, cross-validation is performed on each of these polynomials, and the test MSE is plotted
After the third and forth degree, the test MSE begins to increase  

When using regression splines, using LOOCV, we see that the RSS starts to level off after 6 degrees of freedom, but the test MSE is lowest at 10 degrees of freedom, but the at df =5 or 8, the results were similar

```{r}
# Part C
library(boot)
MSE = rep(NA,10)
for (i in 1:10) {
  model = glm(nox~poly(dis,i),data = Boston)
  MSE[i] = cv.glm(Boston,model,K=10)$delta[1]
}
ggplot() +
  geom_bar(mapping = aes(x = 1:10, y = MSE), stat = "identity")
library(splines)
#Part D, E, and F
MSE = c()
RSS = c()
for (i in 3:20)
{
  
  model = glm(nox~bs(dis, df = i), data = Boston)
  RSS = c(RSS,mean((model$residuals)^2))
  
  # Use K-Fold CV for quicker results
  MSE = c(MSE, cv.glm(Boston,model)$delta[1])
}

plot(3:20,MSE, type = "b", xlab = "df", ylab = "mse")

plot(3:20,RSS, type = "l", xlab = "df", ylab = "rss")

```
