---
title: "College Nonlinear Problem"
output: html_notebook
---
This is Problem 10 from ISLR's Chapter 7 Applied Section

```{r}
library(ISLR)
library(tidyverse)
library(boot)
library(splines)
library(gam)
library(leaps)
set.seed(1)
# Part A - Split Data into Training/Test Set, and then perform forward subset selection to get a subset of predictors
train = sample(1:dim(College)[1],dim(College)[1]/2)
test = -train

regfwd = regsubsets(Outstate~.,data = College[train,], nvmax= 18)
summary(regfwd)
coef(regfwd, id =6)
```
Part B - Fit a GAM on the training data, using out-of-state tuition as
the response and the features selected in the previous step as
the predictors. Plot the results, and explain your findings.

Based off the graphs below, Expend seems very nonlinear with Out-of-State tuition, while Grad.rate looks nonlinear as well, just less so
```{r}
# GAM Model is fit onto the 6 predictors, with the quantitative predictors given 5 degrees of freedom
gam.model = gam(Outstate~Private+s(Room.Board,5)+s(Terminal,5)+s(perc.alumni,5)+s(Expend,5)+s(Grad.Rate,5), data = College, subset= train)
par(mfrow = c(2,3))
plot(gam.model, se = TRUE, col = "blue")


# Part C - Evaluation of the GAM model on the test set, R^2 is outputted
# 77% of the variation of Out of State tuition costs is accounted for by the various predictors in this GAM model
preds = predict(gam.model, newdata = College[test,])
RSS = sum((College[test,]$Outstate - preds)^2)
TSS = sum((College[test,]$Outstate - mean(College[test,]$Outstate))^2)
print(1 - (RSS/TSS))

# Part D
summary(gam.model)
```
Looking at the analysis of variance in the parameters, the Expend parameter has a very strong nonlinear relation, room/board also has a strong nonlinear relationship, while graduation has a moderate nonlinear relationship. Percent.alumni and terminal are weaker nonlinear predictors. Regardless, the analysis of all predictors appear to have a strong effect on the prediction of out of state tuition, as indicated by the p-values of the parametric effects anova
