---
title: "Wage Nonlinear Regression Problem"
output: html_notebook
---
This is Problem 6 from ISLR's Chapter 7 Applied Section 

Problem is below:
"Perform polynomial regression to predict wage using age. Use cross-validation to select the optimal degree d for the polynomial. What degree was chosen, and how does this compare to the results of hypothesis testing using ANOVA? Make a plot of the resulting polynomial fit to the data."

We'll first use the summary function to get a good understanding of how many models to cross validate
Plotting the summary of a 5 degree fit over the entire dataset, we see that up to the fourth degree, the P-Value of our polynomial stays roughly less than or equal to .05. We'll go up beyond this degree for the CV tests
```{r}
# Download Libraries, Attach the Wage Dataset, and get training and test sets
library(ISLR)
library(boot)
library(tidyverse)
attach(Wage)
set.seed(1)
train = sample(3000,2100)
test = -train
fit = lm(wage~poly(age,5), data = Wage)
coef(summary(fit))
detach(Wage)
```

Using K-Fold Cross Validation with K=4, I plotted out the results for degrees 1 through 10, and after degree 3 or 4, things start to level off

```{r}
fits = c()
errors = rep(0,10)
for (i in 1:10) {
 
  fit = glm(wage~poly(age,i), data = Wage)
  fits = c(fits, fit)
  errors[i] = cv.glm(Wage,fit, K= 4)$delta[1]
}
errors

ggplot() + 
  geom_bar(mapping = aes(x = 1:10, y = errors), stat = "identity")
```

Using a step function, we see that CV error is the lowest when the number of cuts is equal to 8
```{r}
errors= rep(NA,9)
for (i in 2:10)
{
  Wage$age.cut = cut(age,i)
  fit = glm(wage ~ age.cut, data = Wage)
  errors[i-1] = cv.glm(Wage,fit,K=4)$delta[1]
}
errors
ggplot() + 
  geom_bar(mapping = aes(x = 2:10, y = errors), stat = "identity")
```

Step Function with 8 cuts is displayed
```{r}
age.range <- range(Wage$age)
age.grid <- seq(from = age.range[1], to = age.range[2])
plot(wage~age, data = Wage, col = "darkgrey")
fit = glm(wage ~ cut(age,8), data = Wage)
pred = predict(fit, list(age = age.grid))
lines(age.grid,pred, col = "red", lwd = 2)
```

