---
title: "Wage other Predictors Problem"
output: html_notebook
---

This is Problem 7 For ISLR's Chapter 7 Applied Section
"The Wage data set contains a number of other features not explored
in this chapter, such as marital status (maritl), job class (jobclass),
and others. Explore the relationships between some of these other
predictors and wage, and use non-linear fitting techniques in order to
fit flexible models to the data. Create plots of the results obtained,
and write a summary of your findings."


We begin by plotting out age and wage, and then coloring in the points based off of their marital status and job class
Based off the data, there is a large cluster of married people in the information jobclass, meaning that this could potentially yield
```{r}
library(ISLR)
library(tidyverse)
library(boot)
set.seed(1)
names(Wage)

ggplot(data = Wage) + 
  geom_point(mapping = aes(x=age,y=wage, col = maritl))
ggplot(data = Wage) + 
  geom_point(mapping = aes(x=age,y=wage, col = jobclass))
summary(Wage$maritl)
summary(Wage$jobclass)


```
We'll be using analysis of variance to see which features to add and which not to add
By analyzing the p-values, combined with the analysis of variance for the fourth predictors, we see that including the marital status and the and jobclass are good predictors for the wage of a given individual, and through our analysis of variance, we see the fourth model, the model with both of those predictors included, did better than the original model without the predictors.

```{r}
library(gam)
library(splines)
library(car)
gam.m1 = gam(wage ~ lo(year, span = .7) + s(age, df = 5) + education, data = Wage)
gam.m2 = gam(wage ~ lo(year, span = .7) + s(age,df=5) + education +jobclass, data = Wage)
gam.m3 = gam(wage ~ lo(year, span = .7) + s(age,df=5) + education +maritl, data = Wage)
gam.m4 = gam(wage ~ lo(year, span = .7) + s(age,df=5) + education +maritl + jobclass, data = Wage)
anova(gam.m1,gam.m2,gam.m3,gam.m4)
summary(gam.m4)
vif(gam.m4)
par(mfrow = c(2, 3))

#plot the best fit
plot(gam.m4, se = T, col = "blue")
```

