---
title: "R Notebook"
output: html_notebook
---



Explores Stock Market Data, it's correlations, and plots the only substantial correlation

```{r}
library(ISLR)
names(Smarket)
dim(Smarket)
summary(Smarket)
cor(Smarket[,-9])
ggplot(data = Smarket, aes(x = Smarket$Year,y = Smarket$Volume)) +
  geom_point()
```

fits a logisitics gression model (glm is a generalized linear model) to predict direction of the stock market given the percentage returns for the five previous trading days, as well as volume
Gets a summary of the glm, the coefficients, and then the p-values

```{r}
glm.fits = glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume,
               data = Smarket, family = binomial)
summary(glm.fits)
summary(glm.fits)$coef
summary(glm.fits)$coef[,4]
```

This prediction predicts the probability of the market going up or down on the training data, and we print out the first 10 probabilities
We also make sure that the probability corresponds to the market going up
Turn these probabilities into predictions with a threshold value of .5
We then output the confusion matrix, and training prediction rate
```{r}
glm.probs = predict(glm.fits,type = "response")
glm.probs[1:10]
contrasts(Smarket$Direction)
glm.pred = rep("Down",1250)
glm.pred[glm.probs > .5] = "Up"
table(glm.pred,Smarket$Direction)
mean(glm.pred == Smarket$Direction)
```

We now use test set data, fit a new model, and use that to predict the dataset
Using the test set to get the error rate from the set
```{r}
train = (Smarket$Year < 2005)
Smarket.2005 = Smarket[!train,]
dim(Smarket.2005)
Direction.2005 = Smarket$Direction[!train]
glm.fit2 = glm(Direction ~ Lag1+Lag2+Lag3+Lag4+Lag5+Volume,
               data = Smarket, family = binomial, subset = train)
glm.probs = predict(glm.fit2,Smarket.2005, type = "response")
glm.pred=rep("Down",252)
glm.pred[glm.probs >.5]="Up"
table(glm.pred ,Direction.2005)
mean(glm.pred != Direction.2005)
```



