---
title: "Linear Regression Lab"
output: html_notebook
---


```{r}
LoadLibraries= function (){
 library(ISLR)
 library(MASS)
 library(tidyverse)
 library(car)
 print("The libraries have been loaded .")
 }

LoadLibraries()
lm.fit = lm(medv~lstat, data = Boston)
summary(lm.fit)
```

We can use the names() function in order to find out what other pieces names() of information are stored in lm.fit
coef() extracts the coefficients, while confint() gets the confidence interals
The predict() function can be used to produce confidence intervals and prediction intervals for the prediction of medv for a given value of lstat
```{r}
names(lm.fit)
coef(lm.fit)
confint(lm.fit)

predict(lm.fit, data.frame(lstat = c(5,10,15)),interval = "confidence")
ggplot(data = Boston) +
  geom_point(mapping = aes(x = lstat, y= medv)) +
  geom_smooth(method = "lm", mapping = aes(x = lstat, y = medv))
ggplot(data = Boston, aes(x = lstat,y= residuals(lm.fit))) +
  geom_point()
```

Does a Multiple Regression Fit, getting the R^2, Relative Standard Error

```{r}
fit = lm(medv~lstat+age, data = Boston)
summary(fit)
fit_all = lm(medv~.,data = Boston)

r_squared = summary(fit_all)$r.sq
RSE = summary(fit_all)$sigma
vif(fit_all)
new_fit =lm(medv???.-age ,data=Boston)
```

Interaction Terms -> beta1:beta2 is the two multiplied together, while beta1*beta2 is the same as beta1+beta2+beta1:beta2
Below also shows a nonlinear fit. Polynomials can also be initialized with poly(lstat,k), corresponding to lstat^k
```{r}
summary(lm(medv~lstat*age, data = Boston))
lm.fit2=lm(medv???lstat+I(lstat^2), data = Boston)
summary(lm.fit2)
anova(lm(medv???lstat, data = Boston),lm.fit2)
```

